\documentclass [hw]{exam}

\usepackage[hw]{physics9}

\title{Practice problems: Momentum and impulse}
\author{\mobeardInstructorShort}
\date{\today}
\duedate{never}

\begin{document}
\maketitle

\begin{questions}
\question A mass $m$ undergoes freefall. What is its linear momentum after it has fallen a distance $h$?

\question While waiting in his car at a stoplight, an \SI{80}{\kilo\gram} man and his car are suddenly accelerated to a speed of \SI{5}{\meter\per\second} as the result of a rear-end collision. Assuming the time taken to be \SI{0.3}{\second}, find the impulse on the man and the average force exerted on him bu the back of the seat of his car. 

\question A camper lets fall a heavy mallet of mass $M$ from a height $y$ upon the top of a tent stake of mass $m$ and drives it into the ground a distance $d$. Find the resistance of the ground, assuming it to be constant and the stake and mallet to stay together on impact. 

\question A golfer hits a ball of mass \SI{51}{\gram} and the ball leaves the club with a velocity of \SI{80}{\meter\per\second}. Assuming that the ball and club are in contact for \SI{0.006}{\second}, find the ball's final momentum and the average force exerted by the club on the ball. 

\question A \SI{5.0}{\gram} bullet moving at \SI{100}{\meter\per\second} strikes a log. Assume that the bullet undergoes uniform deceleration and stops in \SI{6.0}{\centi\meter}. Find the time taken for the bullet to stop, the impulse on the log, and the average force experienced by the log. 

\question A fast neutron has an elastic head-on collision with a proton (hydrogen atom) at rest. Assuming both have the same mass, use the conservation of energy and the conservation of momentum laws to show algebraically what happens.

\question A \SI{1}{\pound} ball moving tot he right at \SI{12}{\foot\per\second} collides head on with a \SI{2}{\pound} ball moving at \SI{12}{\foot\per\second} in the opposite direction. If the two balls stick together after the collision, what is their total momentum after the collision, and their final velocity, including the direction? 

\question An \SI{8}{\gram} bullet is fired horizontally into a \SI{9}{\kilo\gram} block of wood and sticks in it. The block, which is free to move, has a velocity of \SI{40}{\centi\meter\per\second} after impact. Find the initial velocity of the bullet. 

\question A \SI{2}{\kilo\gram} rock is moving at a speed of \SI{6}{\meter\per\second}. What constant force is needed to stop the rock in \SI{7e-4}{\second}?

\question A \SI{20}{\gram} bullet moving horizontally at \SI{50}{\meter\per\second} strikes a \SI{7}{\kilo\gram} block resting on a table. The bullet embeds in the block after collision. Find the speed of the block after collision and the frictional force between the table and block if the block moves \SI{1.5}{\meter} before stopping.

\question A \SI{2500}{\pound} car moving at \SI{55}{mi\per\hour} in the positive direction collides head-on with a  \SI{3500}{\pound} car moving \SI{25}{mi\per\hour} in the negative direction. The two cars stick together after the collision. FInd the momentum (magnitude and direction) of the wreckage immediately after the collision. Also, find the velocity of the cars immediately after the collision.

\question A railroad flatcar of mass $M$ is coasting along a track at speed $v$ when a large machine of mass $m$ topples off a platform and falls straight down onto the car. How fast is the car moving after the machine comes to rest on it? 

\question Prove that relative velocity is reversed by a head-on elastic collision. 
 
\end{questions}
\end{document}