\documentclass[exam,addpoints, noanswers]{exam}

\usepackage[exam]{physics9}

\title{Test \#4}
\date{\printdate{6/2/2021}}
\author{\mobeardInstructorShort}

\begin{document}
\maketitle
\vfill
\mobeardExamNameBlock
\vfill
Instructions: 
\begin{enumerate}
\item Do not open exam until instructor announces that you may begin.
\item Please write your name on each page in case the sheets are separated. 
\item Closed notes, closed book.  You may use two \SI{8.5x11}{\inch} hardcopy note sheets, both sides. You may \textbf{not} use your computer or phone as your notes. 
\item Calculators (TI-84 or equivalent) are permitted.  You may \textbf{not} use your computer or phone as your calculator. 
\item Eight multiple choice are worth five points each, three essay questions (divided into parts) are worth 20 points each. 
\item In answering the essay questions, be thorough but concise. Deep understanding of the concepts will be displayed by proper use of vocabulary and discussion of the interconnectedness of concepts. 
\item Show your calculations and (most importantly) your \textbf{thinking}.
\end{enumerate}
\vfill
\begin{center}
\gradetable[h][questions]
\end{center}
\clearpage



\begin{questions}
\question[5] Easy warmup question: what are correct SI units for momentum? Please select all that apply. 
\begin{choices}
\CorrectChoice \si{\kilo\gram\meter\per\second}
\CorrectChoice \si{\newton\second}
\choice \si{\joule}
\choice \si{slug\ \foot\per\second}
\end{choices}

\question[5] Which of the following are situations in which you might expect momentum to be conserved? Please select all that apply. 
\begin{choices}
\CorrectChoice An inelastic collision between two carts with no outside forces acting
\CorrectChoice An elastic collision between two carts with no outside forces acting
\CorrectChoice An explosion between two carts with no outside forces acting
\choice A yellow Labrador retriever accelerating as he runs to catch a ball
\CorrectChoice An airborne Labrador retriever at the instant of catching a ball, neglecting air resistance. 
\end{choices}

\question[5] The tallest roller coaster at Six Flags Great Adventure is the Kingda Ka (elevation \SI{456}{\foot} (\SI{139}{\meter}), max speed 128 mph (\SI{57}{\meter\per\second}). What physics conservation principle would be most useful in understanding the relationship between elevation $h$ and max speed $v$?
\begin{choices}
\choice conservation of linear momentum
\choice conservation of charge
\choice conservation of mass
\choice conservation or angular momentum
\CorrectChoice conservation of energy
\end{choices}

% sumo wrestlers
\question[5] The Hawaiian Samoan-American \emph{ozeki}-ranked sumo wrestler Konishiki weighed \SI{288}{\kilo\gram} at his peak weight. The next heaviest high ranking sumo wrestler is also a Samoan-American, \emph{yokozuna}-ranked Musashimaru (\SI{237}{\kilo\gram}). During the initial \emph{taich-ai} charge, assume the two wrestlers are both moving at \SI{2}{\meter\per\second} in opposite directions. What is the overall momentum of the system?
\begin{choices}
\choice \SI{1050}{\kilo\gram\meter\per\second}
\CorrectChoice \SI{102}{\kilo\gram\meter\per\second} in the direction Konishiki is traveling
\choice \SI{102}{\kilo\gram\meter\per\second} in the direction Musashimaru is traveling
\choice zero, since they are traveling in opposite directions
\choice Sumo is about technique, not raw strength
\end{choices}

\clearpage
% bocce ball drop calcs
\question[5] A bocce ball is dropped from \SI{2}{\meter} above a device designed to reduce its force on impact. What is the speed of the bocce ball immediately before it hits the device? 
\begin{choices}
\choice \SI{4.4}{\meter\per\second}, using $v=\sqrt{2 g h}$. 
\choice \SI{4.4}{\meter\per\second}, using $v=g t_f$ and $t_f=\sqrt{\frac{2 h}{g}}$. 
\CorrectChoice \SI{6.2}{\meter\per\second}, using either $v=\sqrt{2 g h}$ or $v=g t_f$ and $t_f=\sqrt{\frac{2 h}{g}}$. 
\choice \SI{9.81}{\meter\per\second\squared} as measured in class
\choice \SI{6.2}{\meter\per\second}, using $v=\sqrt{2 g h}$; the other equations $v=g t_f$ and $t_f=\sqrt{\frac{2 h}{g}}$ are not a valid way to compute the speed. 
\end{choices}

% bocce ball drop calcs part two
\question[5] Assume a bocce ball strikes a device designed to reduce its force on impact. The bocce ball ($m=\SI{1.2}{\kilo\gram}$) is initially traveling at \SI{4.4}{\meter\per\second} and comes to rest in \SI{0.1}{\second}. What is the average force the ball encounters? 
\begin{choices}
\CorrectChoice \SI{52.8}{\newton} using $F \Delta t = m \Delta v$. 
\choice \SI{52.8}{\newton} using $F = 0.5 \rho v^2 C_D A$. 
\choice \SI{5.28}{\meter\per\second} using $F \Delta t = m \Delta v$. 
\choice \SI{5.28}{\newton\second} using $F \Delta t = m \Delta v$. 
\choice \SI{5.28}{\kilo\gram\meter\per\second} using $F \Delta t = m \Delta v$. 
\end{choices}

\question[5] In the space below, draw a free body diagram of the bocce ball during its collision with your device.
\begin{solution}[2in]
\end{solution}

\question[5] When a martial artist punches a punching bag, she pivots her hip and pushes off with the ball of her foot. Why?
\begin{choices}
\CorrectChoice By maximizing her momentum ($m\Delta v$) she increases the force she is able to generate on contact with the bag ($F\Delta t$). 
\choice Because it looks cool on Youtube. 
\choice To minimize the recoil on herself, by making her overall momentum close to zero. 
\choice There is no functional role of hip or foot motion. 
\choice When she extends her arm, she must move her hips in order to remain balanced with her center of gravity firmly within the base provided by her stance. 
\end{choices}



\clearpage
% baseball pitching machine converted to hockey use
\question Using what you learned in physics class, you decide to modify a baseball pitching machine to shoot hockey pucks. You plan to market the new devices under the trade name Mo Better Scorer (MBS), and anticipate that with increasing world interest in hockey you are poised to capture a ginormous market share of such training devices. 
\begin{parts}
\part[5] You intend for the machine to launch standard hockey pucks ($m=\SI{170}{\gram}$) at 105 mph (\SI{47}{\meter\per\second}). Calculate the momentum of the puck. 
\begin{solution}[1in]
\end{solution}
\part[5] If the machine and puck were initially stationary, what was the initial momentum of the entire system?
\begin{solution}[1in]
\end{solution}
\part[5] The baseball pitching machine you base your design on weighs \SI{65}{\pound} (\SI{29.5}{\kilo\gram}). Your initial design places the machine on blades for easy movement on the ice. What problems do you foresee? Support your thinking with a particular physics principle and a quick calculation. 
\begin{solution}[2in]
\end{solution}
\part[5] You also wish to have a ``tiebreaker'' mode that fires five pucks in quick succession. Will this make your problem better or worse? Support your thinking with a quick calculation and propose a fix. 
\begin{solution}[2in]
\end{solution}
\end{parts}


\clearpage
% murder victim falling through a trapdoor into a pie machine
\question Using what you learned in physics class, you decide to open a Murderous Barber Shop (MBS) and Pie Factory inspired by the school musical. You believe you could deliver fresher, more flavorful meat pies if you can reduce the processing time, and so you begin concept design of a special meat pie machine installed underneath a barber shop chair with a trap door. 
\begin{parts}
\part[5] The barber shop patrons are to be dropped into the meat pie machine through the trap door. Draw a free body diagram of an unfortunate barber shop patron immediately after falling through the trap door. 
\begin{solution}[1in]
\end{solution}
\part[5] What is the acceleration of the barbershop patron? Assume the MBS and Pie Factory are located on Earth. 
\begin{solution}[1in]
\end{solution}
\part[5] The stack height of the pie machine requires it to be placed \SI{3}{\meter} below the trapdoor. Calculate the speed of the barber shop patrons as they hit the pie machine. 
\begin{solution}[2in]
\end{solution}
\part[5] In testing your device, you find that the force with which barber shop patrons strike the pie machine is unacceptably high. Propose a way to reduce the force with which they strike the pie machine to half. Back up your proposal with a simple design calculation. 
\begin{solution}[2in]
\end{solution}
\end{parts}

\clearpage
% bocce ball hustling
\question Annoyed by continuing damage to the spikeball nets, the administration decides to try to shift attention to a new Mega Bocce Series (MBS), to be fully promoted on ESPN etc as the next big thing. It becomes hugely popular, which has its downsides, as shadier elements seek to profit off it via nefarious schemes. 
\begin{parts}
\part[9] Benny Bitdiddle, a sketchy guy in your advisory, offers to sell you hollowed-out bocce balls that are half the weight of regulation ones. He claims that with the lighter weight, you can bowl it 1.2x faster, and that thus it will be better able to knock other people's bocce balls away from the jack (\emph{pallino}). Benny got a D in physics while you got an A. Is Benny's proposal worth the 50 bucks he is asking for the special ball? Evaluate it with physics principles and equations. 
\begin{solution}[2in]
\end{solution}
\part[9] Dr.~Sketchilista, black market physicist, offers a different sort of ball. He claims that by weighting the ball with lead to be twice the mass of a normal bocce ball, you will better be able to knock the other player's bocce balls; and that your own ball will be less susceptible to being knocked out of the way. Evaluate his claims with physics principles and equations. 
\begin{solution}[2in]
\end{solution}
\part[2] Based on your findings, which ball (if any) would you use to hustle the Mega Bocce Series? 
\begin{solution}[1in]
\end{solution}
\end{parts}

\end{questions}
\end{document}