\documentclass [hw,addpoints,noanswers]{exam}

\usepackage[hw]{physics9}

\title{Quiz: Egg (bocce ball) drop calculations}
\author{\mobeardInstructorShort}
\date{\today}
\duedate{in class}

\begin{document}
\maketitle

\begin{abstract}
Remember to include some simple calculations in your design notebook, and to start reviewing Q3 and Q4 material for the end of year Q4 test! 
\end{abstract}

\begin{questions}
\question[3] We found the acceleration of gravity to be $g=\SI{-9.81}{\meter\per\second}$ when we analyzed the data from dropping things from the balcony. Using this result, compute the speed of the bocce ball after an $x=\SI{1}{\meter}$ drop, at the instant in time $t$ just before it hits your device. You may need the following equations: 
\begin{align*}
x &= \frac{1}{2} a t^2 \\
v &= a t,
\end{align*}
or, alternatively, $GPE=mgh$, $KE=\frac{1}{2}mv^2$.
\begin{solution}[2.5in]
Using energy conservation, $mgh=\frac{1}{2}mv^2$, so $v^2 = 2gh$ or $v=\sqrt{2 g h}=\SI{4.4}{\meter\per\second}.$ Be sure to include proper units. You could also do this problem using kinematics but it will take two steps (solve for $t$ then find $v$). \textbf{It may be smart to include this calculation in your design notebook.}
\end{solution}

\question[2] The mass of the bocce ball is $m=\SI{0.520}{\kilo\gram}$. Assuming $\Delta v=\SI{4.4}{\meter\per\second}$ and the desired average force is \SI{125}{\newton}, what is the minimum allowable $\Delta t$ for the collision? You may need the equation $F \Delta t = m \Delta v$. 
\begin{solution}[2.5in]
Using momentum conservation, $F \Delta t = m \Delta v$, so
\begin{align*}
\Delta t &= \dfrac{m \Delta v}{F} \\
&= \dfrac{\SI{0.520}{\kilo\gram} \cdot \SI{4.4}{\meter\per\second}}{\SI{125}{\kilo\gram\meter\per\second\squared}} \\
&= \SI{0.02}{\second}.\ \text{Check units!} 
\end{align*}
\textbf{It may be smart to include this calculation in your design notebook} and guesstimate $\Delta t$ for a number of design cases. 
\end{solution}
\end{questions}
\end{document}