\documentclass[hw,addpoints,noanswers]{exam}

\usepackage[hw]{../physics9}

\title{(daily) Homework: Energy}
\date{\today}
\duedate{\printdate{4/7/2021}}
\author{\mobeardInstructorShort}

\begin{document}
\maketitle

\begin{abstract}
To help with your problem solving skills and build confidence in applying physics to actual situations we will start having more daily homeworks; this will also help even out the grading. The daily homeworks are meant to be low stress and short; you may use your notes to help recall concepts you may have missed but try your best to answer without lookups. You should not need to spend more than \SI{5}{\min} on this assignment. 
\end{abstract}

\begin{questions}
\question Xuan the Guide Dog Puppy (mass \SI{65}{\pound}, \SI{29.4}{\kilo\gram}) is in the park when he is given the recall command \textbf{come!}. He begins running back towards his handler at \SI{8}{\meter\per\second}. 
\begin{parts}
\part[1] What is the formula for kinetic energy? \answerline[$KE = \frac{1}{2}mv^2$]
\part[1] What is Xuan's mass $m$? \answerline[$m=\SI{29.4}{\kilo\gram}$]
\part[1] What is Xuan's kinetic energy? Please show your work.
\begin{solution}[2in]
\begin{align}
KE &= \frac{1}{2}m v^2 \\
&= 0.5 \cdot \SI{29.4}{\kilo\gram} \cdot (\SI{8}{\meter\per\second})^2\\
&= \SI{941}{\joule}
\end{align}
\end{solution}
%\part Is momentum a scalar or a vector? \answerline[vector]

Xuan encounters a freestyle ski jump ramp that launches him vertically. Assume energy is \textbf{conserved}. 
\part[1] Xuan's kinetic energy is converted to gravitational potential energy. What is the formula for gravitational potential energy, and (assuming perfect conversion) what is Xuan's GPE at the peak of his flight. Please show your work. 
\begin{solution}[2in]
$GPE=mgh$. Xuan's maximum $GPE=\SI{941}{\joule}$. 
\end{solution}
\end{parts}
\end{questions}
\end{document}